---
title: "MachingLearningProject"
author: "Jiao Lai"
date: "3/17/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Step 0.1. Load and aggregate data
```{r}
library(dplyr)
library(tidyverse)
Study.A = read.csv("Study_A.csv")
Study.B = read.csv("Study_B.csv")
Study.C = read.csv("Study_C.csv")
Study.D = read.csv("Study_D.csv")
Study = rbind(Study.A, Study.B, Study.C, Study.D)
Study$P <-
  Study$P1 + Study$P2 + Study$P3 + Study$P4 + Study$P5 + Study$P6 + Study$P7
Study$N <-
  Study$N1 + Study$N2 + Study$N3 + Study$N4 + Study$N5 + Study$N6 + Study$N7
Study$G <-
  Study$G1 + Study$G2 + Study$G3 + Study$G4 + Study$G5 + Study$G6 + Study$G7 + Study$G8 + Study$G9 + Study$G10 + Study$G11 + Study$G12 + Study$G13 + Study$G14 + Study$G15 + Study$G16
Study$Treatment[TxGroup == 'Control'] <- 0
Study$Treatment[TxGroup == 'Treatment'] <- 1
Study$Imp[VisitDay == 0] <- 0
Study$Imp[VisitDay > 0] <- 1
Study$flag[LeadStatus == 'Passed'] <- 0
Study$flag[LeadStatus != 'Passed'] <- 1
Study <- unique(Study)
```
Step 0.2. Drop unpredictive features: SiteID, RaterID, AssessmentiD, TxGroup, Study, LeadStatus.
```{r}
Study <- subset(Study, select = -c(SiteID, RaterID, AssessmentiD, TxGroup, Study, LeadStatus))
Study <- unique(Study)
```

Step 0.3. Check and drop duplicates. 
The duplicates happened because the same patient being assessed or reviewed multiple times on the same day. Since the scores are about the same, we can only keep one of them.

```{r}
Study.dup <-
  Study %>% group_by(PatientID, VisitDay) %>% count() %>% filter(n > 1) %>% select(-n) %>% left_join(Study)
Study.de_dup <-
  Study.dup %>% group_by(PatientID, VisitDay) %>% filter(row_number() == 1)
Study <-
  Study %>% anti_join(Study.de_dup, by = c('PatientID', 'VisitDay'))
Study <- bind_rows(Study, Study.de_dup)

```
Step 0.4. Fix 'Country' error. Check if there is missing data.
```{r}
unique(Study$Country)
Country <-
  Study %>% group_by(PatientID) %>% distinct(Country) %>% filter(Country != 'ERROR')
Country_Error <-
  Study %>% group_by(PatientID) %>% distinct(Country) %>% filter(Country == 'ERROR') %>% left_join(Country, by = 'PatientID')
# Two patients with unknown Country information.
Study$Country[Study$PatientID == '30040'] <- 'USA'
Study <-
  Study[!(Study$PatientID == '30177' | Study$PatientID == '30294'), ]
Study[!complete.cases(Study), ]
```
Step 0.5. Summary data.
```{r}
summary(Study)
plot(Study$PANSS_Total)
plot(Study$P)
plot(Study$N)
plot(Study$G)
```
Objective 1. Detect treatment effect.
Step 1.1. Extract day 0 observations, and last visit day observations.
Only use valid assessment scores.
```{r}
Study.valid <- Study %>% filter(flag == 0)
VisitDay.max = aggregate(VisitDay ~ PatientID, data = Study.valid, max)
Study.Last = merge(Study.valid, VisitDay.max)
Study.Last = Study.Last[Study.Last$VisitDay > 0, ]
Study.Baseline = Study.valid[Study.valid$VisitDay == 0, ]
patient.Last = Study.Last$PatientID
patient.Baseline = Study.Baseline$PatientID
Study.Baseline.TE = Study.Baseline[is.element(Study.Baseline$PatientID, t(patient.Last)), ]
Study.Last.TE = Study.Last[is.element(Study.Last$PatientID, t(patient.Baseline)), ]
Study.TE <-
  Study.Baseline.TE %>% inner_join(Study.Last.TE, by = c('PatientID', 'Country', 'Treatment')) %>% select(
    PatientID,
    Country,
    Treatment,
    VisitDay.y,
    PANSS_Total.x,
    P.x,
    N.x,
    G.x,
    P.y,
    N.y,
    G.y,
    PANSS_Total.y
  ) %>% mutate(diff_T = PANSS_Total.y - PANSS_Total.x) %>% mutate(diff_P = P.y - P.x) %>% mutate(diff_N = N.y - N.x) %>% mutate(diff_G = G.y - G.x) %>% mutate(VisitDay = VisitDay.y) %>% select(PatientID,
                                                                                                                                                                                                 Country,
                                                                                                                                                                                                 Treatment,
                                                                                                                                                                                                 VisitDay,
                                                                                                                                                                                                 diff_T,
                                                                                                                                                                                                 diff_P,
                                                                                                                                                                                                 diff_N,
                                                                                                                                                                                                 diff_G)
Study.TE$PatientID <- as.character(Study.TE$PatientID)

```
Step 1.2. Conduct Linear Regression.
The results show the two groups are comparable, and there is no treatment effect on PANSS_Total, Positive, Negative, or General scores.
```{r}
lm.T = lm(diff_T ~ Treatment + Country, data = Study.TE)
summary(lm.T)
lm.P = lm(diff_P ~ Treatment + Country, data = Study.TE)
summary(lm.P)
lm.N = lm(diff_N ~ Treatment + Country, data = Study.TE)
summary(lm.N)
lm.G = lm(diff_G ~ Treatment + Country, data = Study.TE)
summary(lm.G)
lm.pre.T = lm(PANSS_Total ~ Treatment + Country, data = Study.Baseline.TE)
summary(lm.pre.T)
lm.pre.P = lm(PANSS_Total ~ Treatment + Country, data = Study.Baseline.TE)
summary(lm.pre.P)
lm.pre.N = lm(PANSS_Total ~ Treatment + Country, data = Study.Baseline.TE)
summary(lm.pre.N)
lm.pre.G = lm(PANSS_Total ~ Treatment + Country, data = Study.Baseline.TE)
summary(lm.pre.G)
lm.T.FilterByVisitDay = lm(diff_T ~ Treatment + Country, data = Study.TE[Study.TE$VisitDay > 120, ])
summary(lm.T.FilterByVisitDay)
lm.P.FilterByVisitDay = lm(diff_P ~ Treatment + Country, data = Study.TE[Study.TE$VisitDay > 120, ])
summary(lm.P.FilterByVisitDay)
lm.N.FilterByVisitDay = lm(diff_N ~ Treatment + Country, data = Study.TE[Study.TE$VisitDay > 120, ])
summary(lm.N.FilterByVisitDay)
lm.G.FilterByVisitDay = lm(diff_G ~ Treatment + Country, data = Study.TE[Study.TE$VisitDay > 120, ])
summary(lm.G.FilterByVisitDay)
```


Objective 2. Patients segmentation.
Step 2.1. Data pre-processing. Only use day 0 observations. Drop outliers.
```{r}
Study.Baseline = Study[Study$VisitDay == 0,]
plot(Study.Baseline$PANSS_Total)
plot(Study.Baseline$P)
plot(Study.Baseline$N)
plot(Study.Baseline$G)
Study.Baseline = Study.Baseline[Study.Baseline$PANSS_Total<=120,]
Study.Baseline = Study.Baseline[Study.Baseline$PANSS_Total>=60,]
Study.Baseline = Study.Baseline[Study.Baseline$G>=25,]
```
Step 2.2. k-means clustering
```{r}
set.seed(3)
km.out1 = kmeans(Study.Baseline[,4:33], 2, nstart = 25, iter.max = 500)
km.out1
km.out2 = kmeans(Study.Baseline[,4:33], 3, nstart = 25, iter.max = 500)
km.out2
km.out3 = kmeans(Study.Baseline[,4:33], 4, nstart = 25, iter.max = 500)
km.out3
km.out4 = kmeans(Study.Baseline[,4:33], 5, nstart = 25, iter.max = 500)
km.out4
```
Step 2.3. Select optimal k = 3.
```{r}
km.out = kmeans(Study.Baseline[, 4:33], 3, nstart = 25, iter.max = 500)
km.out
cluster = km.out$cluster
result = cbind(Study.Baseline, cluster)
library(scatterD3)
scatterD3(
  data = result,
  x = P,
  y = N,
  col_var = cluster,
  size_var = G
)
```
Goal 3. Last visit assessment score prediction.
Step 3.1. Data pre-processing. Select last visit assessment, and second last visit assessment.
```{r}
G3.Last <-
  Study.Last %>% select(PatientID, Country, VisitDay, PANSS_Total)
G3._Last <-
  Study.valid %>% anti_join(G3.Last, by = c('PatientID', 'VisitDay'))
VisitDay.2ndmax = aggregate(VisitDay ~ PatientID, data = G3._Last, max)
G3.2ndLast = merge(G3._Last, VisitDay.2ndmax)
G3.Last <-
  G3.Last %>% inner_join(G3.2ndLast, by = c('PatientID', 'Country')) %>% select(PatientID,
                                                                                Country,
                                                                                PANSS_Total.x,
                                                                                PANSS_Total.y,
                                                                                VisitDay.x,
                                                                                VisitDay.y)

```
Step 3.2. Conduct linear regression.
The results show that most recent score is highly predictive for last visit assessment scores.
```{r}
G3.v1 <-
  G3.Last %>% filter(VisitDay.x - VisitDay.y <= 21) %>% filter(VisitDay.x > 120)
lm.G3.v1 <-
  lm(PANSS_Total.x ~ PANSS_Total.y + Country, data = G3.v1)
summary(lm.G3.v1)
G3.v2 <-
  G3.Last %>% filter(VisitDay.x - VisitDay.y <= 14) %>% filter(VisitDay.x > 120)
lm.G3.v2 <-
  lm(PANSS_Total.x ~ PANSS_Total.y + Country, data = G3.v2)
summary(lm.G3.v2)
G3.v3 <-
  G3.Last %>% filter(VisitDay.x - VisitDay.y <= 28) %>% filter(VisitDay.x > 120)
lm.G3.v3 <-
  lm(PANSS_Total.x ~ PANSS_Total.y + Country, data = G3.v3)
summary(lm.G3.v3)
```
Step 3.3. Make forecasting.
```{r}
forecast = read.csv("Study_E.csv")
# Apply linear regression on study.
lm.G3.v3 <-
  lm(PANSS_Total.x ~ PANSS_Total.y + Country, data = G3.v3)
summary(lm.G3.v3)
lm.pred = predict(lm.G3.v3, forecast)
forecast$forecast_pred <- lm.pred
```
Goal 4. Classification
Step 4.1. Data pre-processing.
All observations are used for this task.
Create column 'relavant_visit_day'.
```{r}
Study <-
  Study %>% arrange(PatientID, VisitDay) %>% group_by(PatientID) %>% mutate(relevant_visit_day = row_number())
Study
Study_mean_byGroup <-
  Study %>% select(VisitDay, PANSS_Total, P, N, G, relevant_visit_day, flag) %>% group_by(flag) %>% summarise(
    mean_VisitDay = mean(VisitDay),
    mean_PANSS_Total = mean(PANSS_Total),
    mean_P = mean(P),
    mean_N = mean(N),
    mean_G = mean(G),
    mean_relevant_visit_day = mean(relevant_visit_day)
  )
Study_var_byGroup <-
  Study %>% select(VisitDay, PANSS_Total, P, N, G, relevant_visit_day, flag) %>% group_by(flag) %>% summarise(
    var_VisitDay = var(VisitDay),
    var_PANSS_Total = var(PANSS_Total),
    var_P = var(P),
    var_N = var(N),
    var_G = var(G),
    var_relevant_visit_day = var(relevant_visit_day)
  )

```
Step 4.2. Build logistic regression.
```{r}
set.seed(123)
train = sample(nrow(Study),
               0.8 * nrow(Study),
               replace = FALSE,
               prob = NULL)
test = -train
Study.train = Study[train,]
Study.test = Study[test,]
glm.fit = glm(
  flag ~ PANSS_Total + VisitDay + relevant_visit_day  + P + N + Treatment,
  data = Study.train,
  family = 'binomial'
)
summary(glm.fit)
test.pred.prob = predict(glm.fit, Study.test, type = "response")
test.pred <- ifelse(test.pred.prob > 0.5, 1, 0)
mean(test.pred != Study.test$flag)
table(test.pred, Study.test$flag)
train.pred.prob = predict(glm.fit, Study.train, type = "response")
train.pred <- ifelse(train.pred.prob > 0.5, 1, 0)
mean(train.pred != Study.train$flag)
```
Step 4.3. Find optimal threshold.
```{r}
test.pred.prob = predict(glm.fit, Study.test, type = "response")
thresholds = seq(0, 1, 0.01)
best_accuracy = 0
best_threshold = 0
for (threshold in thresholds)
{
  test.pred <- ifelse(test.pred.prob > threshold, 1, 0)
  accu = mean(test.pred == Study.test$flag)
  if (accu > best_accuracy)
  {
    best_accuracy = accu
    best_threshold = threshold
  }
}

```
Step 4.4. Make prediction.
```{r}
Study.E = read.csv("Study_E.csv")
Study.E <-
  Study.E %>% arrange(PatientID, VisitDay) %>% group_by(PatientID) %>% mutate(relevant_visit_day = row_number())
Study.E$P <-
  Study.E$P1 + Study.E$P2 + Study.E$P3 + Study.E$P4 + Study.E$P5 + Study.E$P6 + Study.E$P7
Study.E$N <-
  Study.E$N1 + Study.E$N2 + Study.E$N3 + Study.E$N4 + Study.E$N5 + Study.E$N6 + Study.E$N7
Study.E$G <-
  Study.E$G1 + Study.E$G2 + Study.E$G3 + Study.E$G4 + Study.E$G5 + Study.E$G6 + Study.E$G7 + Study.E$G8 + Study.E$G9 + Study.E$G10 + Study.E$G11 + Study.E$G12 + Study.E$G13 + Study.E$G14 + Study.E$G15 + Study.E$G16
Study.E$Treatment <- ifelse(Study.E$TxGroup == 'Treatment', 1, 0)
E.pred.prob = predict(glm.fit, Study.E, type = "response")
Study.E$flag <- ifelse(E.pred.prob > best_threshold, 1, 0)

```